\section{Introduction}
When developing a web application, there are numerous issues the application writer must consider. One of the most used languages when designing the front-end of a web application is JavaScript\cite{javascript_popularity}, a dynamically typed, multi-paradigm programming language.\cite{javascript_info} Unfortunately, JavaScript suffers from a couple of drawbacks and attacks towards the language through Cross-Site Scripting (XSS) were on the third place on the OWASP Top 10 in 2013.\cite{owasp_xss_rank} In a XSS attack, the attacker manages to inject JavaScript code into websites that are considered secure.\cite{owasp_xss, excess_xss} These injections can do anything from harmless pranks (e.g. showing an alert box) to redirecting the non-suspecting user to a fake website to steal valuable information. When an attack like XSS succeeds, the culprit is usually non-escaped input from the users. If a website directly includes the user input, an attacker can insert input that will be treated as code by the victim.\cite{excess_xss}
\subsection{Problems with JavaScript}
Apabepacepa
\subsection{Dynamic Typing}
Trololol
\subsubsection{Examples of attempts to solutions}
There have been several attempts of providing more security to JavaScript, everything from creating a statically typed language that compiles to JavaScript to a compiler from an already existing programming language and compile it to JavaScript to a static type checker. Examples of attempted soultions are
\begin{itemize}
  \item \textbf{TypeScript}\cite{typescript}, a typed superset of JavaScript that compiles to plain JavaScript.
  \item \textbf{TeJaS}\cite{tejas-art,tejas-git}, which allows you to annotate type signatures as comments in the JavaScript code and then type checks the code.
  \item \textbf{GHCJS}\cite{ghcjs} and \textbf{Haste}\cite{haste-lang,haste-symposium}, which compiles from Haskell, a statically typed, high-level functional programming language\cite{haskell}, to JavaScript.
\end{itemize}

\subsection{Information-Flow Control}
Talk about IFC...
\subsubsection{JSFlow}
JSFlow is awesome!
